Truthfulness Verification System

### 2011-04-13 proposal/paper.tex
checked updade with totha
Commitiing till last update

@@

After getting the wikipedia page we can convert the html file into text file and then we can separate all the sentences in different lines. The sentence is the part of the whole text that exist between two full-stops. Another important observation is that the relevant information will be within the same sentence in the question-answer type sentences almost all the time. So we can match the content words to the sentences extracted from Wikipedia page and verify the truthfulness of the sentence.

@@

When we do a regular expression search of the $DU$ "Les Paul" in these two sentences, we do not get any match. Hence we can conclude that the statement is not true.

@@

\begin{itemize}
\item 
\end{itemize}
\section{Conclusion}
It this paper, we described current system for truthfulness verification of statements found in the web called T-Verifier and described the possible improvements to the current system. We will primarily use authentic information sources such as Wikipedia and Ontologies extracted from Wikipedia such as FreeBase, DBPedia, and YOGA to found the truthfulness of doubtful statements.

>>>>

{\em Mar 29 to Apr 12}
See page~[\pageref{Mar20toApr12}] of section~\ref{sec:UseOfWikipedia}
\begin{itemize}
\item Extraction of meaningful entities(noun phrase) in topic unit and dount unit by checking all combination of content words
\item Ranking based on co-occurence of content words in the interesction of bag of words generated from wiki-articles. 
\item Extracting 
\end{itemize}

@@

%When we do a regular expression search of the $DU$ "Les Paul" in these two sentences, we do not get any match. Hence we can conclude that the statement is not true.

\label{Mar29toApr12}

\subsubsection{Bag of words}
\begin{enumerate}
\item Step I: Detection of wiki articles from $DU$, $AU$, $TU$.
\begin {enumerate}
\item For each of them we use a basic exponential algorithm that takes all possible combinations of the content words.
\item The default configuration of T-verifier generates a total of 5 $AU$s, without any ranking.  For each content wordin a $AU$, or a combination of words in case the $AU$ consists of multiple content words, we check if there exists a wikipedia article. There may be three cases:
\begin{enumerate}
\item No article. Continue with next word or word combination. 
\item One article. Go to next step
\item More than one article. The page lands into a "disambiguation" page. We discuss this later. 
\end{enumerate}
\item Similarly for $DU$ we check all possible combinations.
\end{enumerate}
\item We preporcess the wikiaticles by stripping them of html, and converting them into simple bag of words i.e we retain word repetation.
\item We select the following pairwise combination - a wikiarticle detected from $AU$, ($W_{au_{i}}$) and a wikiarticle detected from $TU$ ($W_{tu_{j}}$).
\item We take the common words $c_{ij} = {w_{1}, w_{2}, \ldots}$ matched by regular expressions in these two bags.
\item Since this is a regular expression matching we will have two different values for $c_{ij}$ = common($W_{au_{i}}$, $W_{tu_{j}}$) and $c_{ji}$ = common($W_{tu_{j}}$, $W_{au_{i}}$). This is explained below.
\item In this collection of common words, $c_{ij}$, we search each of the content words of the $TU$, not included in the title of $W_{tu_{j}}$ 
\item We observe the maximum number of matches in all of such interesctions and use the count to rank the $AU$-s generated by T-verifer. 
\end{enumerate}

\paragraph{Handling Disambiguation}
For the testing that we have done with the five problematic sentences, we manually first used the disambiguation page itself in the above process. But as expected, it did not yeild good results. We are experimenting with the following ideas for disabmigutation.
\begin{enumerate}
\item Apply the above steps for each of the page linked from the disambiguation page and get the one which gives the maximum count
\item Use the heading, one sentence summary in the disambiguation page adjacent to the page name, and categories tagged to that article page for each of the disambiguation item.
\end{enumerate}
For this report, we manually did the disambiguation and found the results were comparitively better.

\paragraph{Word matching}
We found that the intersection of the two bags of words, $W_{au_{i}}$ and $W_{tu_{i}}$ produced different results. This is because the regular expression matcher we are using, grep finds the maximal match. Lets consider that the first file, $W_{au_{i}}$ has "invent", and the second file, $W_{tu_{i}}$ has "invented", and "inventor". So when we pick a word in $W_{au_{i}}$ and try to match words in $W_{tu_{i}}$, we will find 2 matches. However, the words from second file, both "invented", and "inventor" will not find any matches in ($W_{au_{i}}$). Taking both way intersections will ensure we do not miss out different morphological forms of the same word appearing in the context common to the both articles.

\paragraph{Example}

A description of the above steps applied to the five problematic sentences
\begin{enumerate}
\item Sentence 1: Les Paul invented the electric guitar.
\begin{enumerate}
\item AU: Les\_Paul, Gibson Les, Leo, Llyod, Adolph Rickenbacker. 
\item Wiki articles detected from AU($W_{au_{i}}$): Les\_Paul, Leo (disambiguation page), Adolph\_Rickenbacker 
\item Wiki articles detected from TU($W_{tu_{i}}$): The\_guitar, The\_Electric\_guitar  Electric\_guitar, invented \rightarrow Invention 
\end{enumerate}
\end{enumerate}

\subsubsection{Sentence Splitting}
After getting the wikipedia page we can convert the html file into text file and then we can separate all the sentences in different lines. The sentence is the part of the whole text that exist between two full-stops. Another important observation is that the relevant information will be within the same sentence in the question-answer type sentences almost all the time. So we can match the content words to the sentences extracted from Wikipedia page and verify the truthfulness of the sentence.

@@

@@

\section{Ranking Algorithm 2}
The main idea behind this ranking algorithm for alternative units is that Wikipedia should contain content words from doubt statement in the same sentence because of the similarity in the context. Details description of the algorithm is given below.

\begin{algorithm}
\small
\caption{Ranking Algorithm 2}
\label{alg:AssignScore} 
\begin{algorithmic}
\STATE Search google with topic unit
\STATE take first Wikipedia page from google search result
\STATE Extract all sentences from the Wikipedia page
\FOR{each alternate unit}
	\FOR{each sentence from Wikipedia page}
		\IF{alternate unit is found in the sentence}
			\IF{some content word found in the sentence}
				\STATE return true
			\ELSE
				\STATE return false	
			\ENDIF
		\ENDIF
	\ENDFOR
\ENDFOR		
\STATE assign same score to all true matches

\end{algorithmic}
\end{algorithm}

\subsection{Examples}
Examples of the algorithm described above is discussed below. The algorithm 2 produce perfect result for three of the five sentences. But it struggles with the remaining two sentences. However, the algorithm 2 primarily augment the algorithm 1 and add strength to the results from algorithm 1. Hence they work better together and produce overall better output.

\subsubsection{Tom Hanks was the lead actress  int move 'Sleepless in Seattle'}
The alternative units for this sentence are "Tom Hanks", "Meg Ryan", "Hanks Meg", "Nora", and "Sam". Here the sentence containing "Tom Hanks" contain the word actor but not actress. So the algorithm classify it as false. But the sentence containing "Meg Ryan" has the word actress and hence classify it as true. For other alternative units, we don't even get any sentences containing the alternative unit. So the algorithm produce an unambiguous single true statement for this sentence.

\subsubsection{Les Paul invented the electric guitar}
The alternative units for this sentence are "Les Paul", "rickenbacker", "Gibson", "Leo", and "Lloyd". Now there is inherent ambiguity about the true inventor of the electric guitar. And Wikipedia does not specify rickenbacker as the inventor of electric guitar. Rather it specify Les Paul as the inventor in several places. Our algorithm unambigiously classify Les Paul as inventor of electric guitar in a similar way for the Sleepless in Seattle case described in the previous subsection.

\subsubsection{Luna 2 is the first spacecraft on the moon}
The alternative units for this sentence are "Luna 2", "Apollo", "Land", "Landing", and "Nasa". Our algorithm returns true for both Luna 2 and Apollo and return false for all others. Now, both Luna 2 and Apollo are in reality true considering the facts that Luna 2 was the first unmanned spacecraft and Apollo was the first manned spacecraft. Hence, we can't say one of the Luna 2 or Apollo as false if farther information is not given. So algorithm 2 can produce correct result for this sentence also.

\subsubsection{English is the Primary Language of the Philippines}
The alternative units for this sentence are "English", "Filipino", "Spanish", "Education", and "History". Our current algorithm struggle with this sentence because of occurrence of content words and alternative units in the same sentence for all the alternative units. So it classify all of them as true and assign same weight to all of them. However, as our algorithm 1 can rank the alternative units in correct order, merged rank from algorithm 1 and algorithm 2 is still correct and produce overall correct result.

\subsubsection{Sullivan is the fastest swimmer in the world}
The alternative units for this sentence are "Sullivan", "Michael phelps", "ALexander", "Long", and "Ernest". Similar to the previous sentence, the algorithm 2 can not produce good ranking for for this sentence. However, our algorithm 1 produces correct ranking and together they produce overall correct result

\section{Conclusion}
It this paper, we described current system for truthfulness verification of statements found in the web called T-Verifier and described the possible improvements to the current system. We will primarily use authentic information sources such as Wikipedia and Ontologies extracted from Wikipedia such as FreeBase, DBPedia, and YOGA to found the truthfulness of doubtful statements.
